---
title: "Web Scraping"
author: "AnnaClaire Marley"
date: "6/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(httr)
library(raster)
library(lubridate)

```

This Rmd scrapes data from HYCOM for salinity and SST in the Gulf of Mexico. Data available from HYCOM is at 3 hour time intervals daily for each year. This code:

* Allows your to choose which year of data to extract, the months you want to extract and which hours of the day to extract
* Creates average monthly minimum and maximum SST and salinity rasters for each year by taking the average of daily minimums and maximums for each month

#### HYCOM Gulf of Mexico 

Scrape netcdf data from "https://www.hycom.org/data/gomu0pt04/expt-50pt1"

```{r}

year_range <- c(2011) # set the year you want to look at
month_range <- c(1:12) # set the months
hours = c(0, 3, 6, 9, 12, 15, 18, 21) # set the hours


for (year in year_range){
  
  for (month in month_range){
    
    # get the number of days in each month
    num_days = days_in_month(as.Date(paste0(year, "-", month, "-01")))
    
    # make vectors
    month_temp_min = c()
    month_temp_max = c()
    month_sal_min = c()
    month_sal_max = c()
    
    for (day in 1:num_days) {
      
      day_data_temp = c()
      
      day_data_sal = c()
      
      for (hour in hours) {
        
        file_name = sprintf("hycom_gomu_501_%d%02d%02d00_t0%02d.nc", year, month, day, hour)
       
        # generate url
        url = sprintf("http://ncss.hycom.org/thredds/ncss/datasets/GOMu0.04/expt_50.1/data/netcdf/%d/%s?var=salinity&var=water_temp&disableLLSubset=on&disableProjSubset=on&horizStride=1&time_start=%d-%02d-%02dT%02d%%3A00%%3A00Z&time_end=%d-%02d-%02dT%02d%%3A00%%3A00Z&timeStride=1&vertCoord=&accept=netcdf", 
                      year, file_name, year, month, day, hour, year, month, day, hour) 
        
        # response
        r <- GET(url)
        
        # check if any errors, want status code to be 200
        print(sprintf("File: %s downloaded. Status code: %d", file_name, status_code(r)))
        
        if (status_code(r) == 200) {
        
        # save file
        bin <- content(r, "raw")
        writeBin(bin, sprintf("../data/HYCOM/Year_%d/%s", year, file_name))
        
        # extract sea surface temp and salinity
        sst <- brick(sprintf("../data/HYCOM/Year_%d/%s", year, file_name), varname = "water_temp")
        salinity <- brick(sprintf("../data/HYCOM/Year_%d/%s", year, file_name), varname = "salinity")
        
        # append to vector
        day_data_temp = c(day_data_temp, sst)
        day_data_sal = c(day_data_sal, salinity) 
        
        
       
        }
      }
    
      # stack the days  
     day_temp_stack <- raster::stack(day_data_temp)
     
     day_sal_stack <- raster::stack(day_data_sal)
      
     # get min and max raster
     temp_min <- min(day_temp_stack)
     temp_max <- max(day_temp_stack)
     sal_min <- min(day_sal_stack)
     sal_max <- max(day_sal_stack)
     
     # append
     month_temp_min = c(month_temp_min, temp_min)
     month_temp_max = c(month_temp_max, temp_max)
    month_sal_max = c(month_sal_max, sal_max)
     month_sal_min = c(month_sal_min, sal_min)
     
     # delete files
    
     folder = sprintf("../data/HYCOM/Year_%d/", year)
     
     for (file in list.files(folder, pattern = "*.nc$")) {
       
       file.remove(paste0(folder, file))
       
     }
     
    
    }  
    
    # save files, average monthly rasters
    month_av_temp_min <- mean(stack(month_temp_min))
    writeRaster(month_av_temp_min, 
                sprintf("../data/HYCOM/Year_%d/av_sst_min_%d-%02d.tif", 
                        year, year, month))
    
    month_av_temp_max <- mean(stack(month_temp_max))
    writeRaster(month_av_temp_max, 
                sprintf("../data/HYCOM/Year_%d/av_sst_max_%d-%02d.tif", 
                        year, year, month))
    
    month_av_sal_max <- mean(stack(month_sal_max))
    writeRaster(month_av_sal_max, 
                sprintf("../data/HYCOM/Year_%d/av_sal_max_%d-%02d.tif", 
                        year, year, month))
    
     month_av_sal_min <- mean(stack(month_sal_min))
    writeRaster(month_av_sal_min, 
                sprintf("../data/HYCOM/Year_%d/av_sal_min_%d-%02d.tif", 
                        year, year, month))
    
    
  
  }
  
  
}
print("finish")

```


```{r}

x <- raster("../data/HYCOM/av_sst_min_2012-01.tif")

plot(x)

```



```{r}
# minimum temperature for each day
# average minimum monthly temperature



# for both water temp and sailinty 
brick("../data/hycom_gomu_501_2012010100_t000.nc", varname = "water_temp")

# then group by day and average for each day then average by month and save that



```

