---
title: "HYCOM Suitability"
author: "AnnaClaire Marley"
date: "6/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#load packages
library(tidyverse)
library(raster)
library(sdmpredictors)
```


# Suitability 

**General Seaweed:**
- Depth: 10-200m
- N:P ratio: between 4 and 80
-----------------------------------------
- Sea surface temperature range: 0-35°C
- Current velocity: 0.1-1 m/s

*Sargassum* specific:
- Optimal sea surface temp: 18-30°C
- Optimal salinity: 24-42 psu

*Euchema* specific:
- Optimal sea surface temp: 25-30°C
- Optimal salinity: 28-34 psu

*Gracilaria* specific:
- Optimal sea surface temp: 15-30°C
- Optimal salinity: 15-25 psu

*Gracilaria spp* from Lisa specific:
- Optimal sea surface temp: 20-30°C
- Optimal salinity: 25-40 psu
- Current velocity: 0.05-0.5 m/s

**Seaweed Parameters**
```{r}
# depth
min_depth <- -10
max_depth <- -200

# Nitrate:Phosphate ratio
n_p_ratio_min <- 4
n_p_ratio_max <- 80

# sea surface temperature
min_sst_value <- 20
max_sst_value <- 30

# salinity
min_sal_value <- 25
max_sal_value <- 40

# current velocity
min_curvel_value <- 0.1
max_curvel_value <- 1

```

```{r}
# set up path to load activate layers
path <- "../data/Active_Data_Layers/"
```


Depth
```{r}

# read in raster
depth_mask <- raster(paste0(path,"/depth_mask_hy.tif"))

# Reclassification matrix 
rcl_matrix_depth <- c(-Inf, max_depth, 0,
                   max_depth, min_depth, 1,
                   min_depth, 0, 0,
                   0, Inf, 0)

# Reclassify 
depth_binary <- reclassify(depth_mask, rcl = rcl_matrix_depth)

# Plot reclassified layer
plot(depth_binary)

```

```{r}
# load function to relcassify rasters
source('../functions/reclass_data.R')
```

Max SST
```{r}
# read in raster
max_sst <- raster(paste0(path,"/max_sst_hy_an_mask.tif"))

# reclassify
max_sst_binary <- reclass_data(raster = max_sst, max_thresh = max_sst_value, 
                        one_thresh = TRUE, which_thresh = "MAX")


# plot reclassified layer
plot(max_sst_binary)

```


Min SST
```{r}
# read in raster
min_sst <- raster(paste0(path,"/min_sst_hy_an_mask.tif"))

plot(min_sst)

# reclassify
min_sst_binary <- reclass_data(raster = min_sst, min_thresh = min_sst_value, 
                        one_thresh = TRUE, which_thresh = "MIN")


# plot reclassified layer
plot(min_sst_binary)

```

Salinity
```{r}
# read in files
sal_max <- raster(paste0(path,"/min_sal_hy_an_mask.tif"))
sal_min <- raster(paste0(path,"/min_sal_hy_an_mask.tif"))


# reclassify
sal_max_binary <- reclass_data(raster = sal_max, max_thresh = max_sal_value, 
                        one_thresh = TRUE, which_thresh = "MAX")
sal_min_binary <- reclass_data(raster = sal_min, min_thresh = min_sal_value, 
                        one_thresh = TRUE, which_thresh = "MIN")

# plot
plot(sal_max_binary)
plot(sal_min_binary)

```

N:P ratio
```{r}
# read in raster
n_p_ratio <- raster(paste0(path,"/n_p_ratio.tif"))

# resample
n_p_ratio_resample_hy <- resample(n_p_ratio, sal_max, method='ngb', progress='text')

# reclassify
np_binary <- reclass_data(raster = n_p_ratio_resample_hy, min_thresh = n_p_ratio_min,
                          max_thresh = n_p_ratio_max, one_thresh = FALSE)

# plot reclassified layer
plot(np_binary)
```





### Final Maps

```{r}
# source functions to save maps created
source("../functions/save_map.R")
```

Suitability areas -- all bio-oracle
```{r}

suitable_areas <- raster::overlay(np_binary,
                                  depth_binary, 
                                  max_sst_binary,
                                  min_sst_binary, 
                                  sal_max_binary, 
                                  sal_min_binary,
                          fun = function(a, b, c, d, e, f){a*b*c*d*e*f})
# zoom in to see
suitable_zoom <- crop(suitable_areas, extent(-100, -80, 22, 35))


# plot
plot(suitable_zoom)
map('world',fill=T,add=T,col='gray')


```

```{r}
### save the map ###
#save_map(png_name = "seaweed_gen_suitable.png", title = "General Seaweed Suitability", 
#         color = "light blue")

# sargassum
save_map(png_name = "sargassum_suitable.png", title = "Sargassum Suitability", 
         color = "purple")
# Euchema
#save_map(png_name = "euchema_suitable.png", title = "Euchema Suitability", 
#         color = "green")
# Gracilaria
#save_map(png_name = "gracilaria_suitable_L.png", title = "Gracilaria Suitability", 
 #        color = "blue")
```

# Exclusion Layers

```{r}
# Marine protected areas
mpa <- raster(paste0(path,"/mpas_binary_hy.tif"))

# Shipping lanes
shipping_lanes <- raster(paste0(path,"/ship_binary_hy.tif"))

# Submarine Cable Areas
sub_cable_areas <- raster(paste0(path,"/sub_cable_area_binary_hy.tif"))

# Pipeline Areas
pipeline_areas <- raster(paste0(path,"/pipe_binary_hy.tif"))

# Military danger zones
military_danger <- raster(paste0(path,"/danger_binary_hy.tif"))

# Oil and Gas platforms
oil_gas_platforms <- raster(paste0(path,"/og_binary_hy.tif"))

```

# Final Suitability Map
```{r}

suitable_areas <- overlay(np_binary, 
                          depth_binary, 
                          max_sst_binary,
                          min_sst_binary, 
                          sal_max_binary, 
                          sal_min_binary,
                          mpa,
                          shipping_lanes,
                          sub_cable_areas,
                          pipeline_areas,
                          military_danger,
                          oil_gas_platforms,
                          fun = function(a, b, c, d, e, f, g, h, i, j,  k, l){a*b*c*d*e*f*g*h*i*j*k*l})

plot(suitable_areas)
map('world',fill=T,add=T,col='gray')

# zoom in to see
suitability_zoom <- crop(suitable_areas, extent(-100, -80, 22, 35))

plot(suitability_zoom)
map('world',fill=T,add=T,col='gray')

```